---
title: "Mammogram"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(survey)
library(car)
library(jtools)
```

Read in the data.
```{r, import, warning = FALSE, message = FALSE}
cancer = read_csv("./data/cancerxx.csv") %>%
  select(HHX, FMX, FPX, WTFA_SA, STRAT_P, PSU_P, REGION, MAMHAD, MAM6YR, 
         RMAM1_MT, RMAM1YR, RMAM1N, RMAM1T, RMAM2, RMAM3A, RMAM3B, MAMPAY, 
         MAMREAS, MDRECMAM, MAMDNBR, MAMABN1, MFOLLOW0, MFOLLO01, MFOLLO02, 
         MFOLLO03, MFOLLO04, MFOLLO05, MNOTFOL1, MAMMODE, MAMCAN1)

adult = read_csv("./data/samadult.csv") %>%
  select(HHX, FMX, FPX, AUSUALPL, AHCPLROU, AHCPLKND, FLA1AR)

family = read_csv("./data/familyxx.csv") %>%
  select(HHX, FMX, RAT_CAT4, RAT_CAT5)

person = read_csv("./data/personsx.csv") %>%
  select(HHX, FMX, FPX, AGE_P, EDUC1, SEX, NOTCOV, COVER65, COVER65O, LA1AR,
         LCONDRT, LACHRONR, HISCODI3, RACRECI3, COVER)

mam_dat = cancer %>%
  left_join(adult, by = c("HHX", "FMX", "FPX")) %>%
  left_join(person, by = c("HHX", "FMX", "FPX")) %>%
  left_join(family, by = c("HHX", "FMX")) 
```

Data Manipulation
```{r}
# outcome is having a mammogram in the last 2 years:RMAM3A = 1,2
mam_dat = mam_dat %>%
  mutate(mam_2 = if_else(RMAM3A <= 2, 1, 0))

# create the age category
mam_dat = mam_dat %>% 
  mutate(age_cat = case_when(AGE_P >= 25 & AGE_P < 40 ~ "25–39",
                             AGE_P >= 40 & AGE_P < 50 ~ "40–49",
                             AGE_P >= 50 & AGE_P < 65 ~ "50–64",
                             AGE_P >= 65 ~ "65+"))
# create educ category
mam_dat = mam_dat %>% 
  mutate(educ_cat = case_when(EDUC1 < 13 ~ "Less than high school",
                              EDUC1 >= 13 & EDUC1 < 15 ~ "High school",
                              EDUC1 >= 15 & EDUC1 < 18 ~ "Some college",
                              EDUC1 >= 18 & EDUC1 <= 21 ~ "College graduate"))

# create financial category
mam_dat = mam_dat %>% 
  mutate(finc_cat = case_when(RAT_CAT5 <= 7 |  RAT_CAT5 %in% c(15, 16) ~ "<200%",
                              RAT_CAT5 %in% c(8, 9) ~ "200–299%", 
                              RAT_CAT5 %in% c(10, 11) ~ "300–399%",
                              RAT_CAT5 >= 18 & EDUC1 <= 21 ~ "400–499%",
                              RAT_CAT5 == 14  ~">=500%",
                              RAT_CAT5 == 17  ~">=200%, no further detail",
                              RAT_CAT5 %in% c(96, 99) ~ "Unknown"))

# create as usual category
mam_dat = mam_dat %>% 
  mutate(ausualpl_cat  = case_when(AUSUALPL == 2 ~ "No",
                                   AUSUALPL %in% c(1, 3) ~ "Yes",
                                   AUSUALPL %in% c(7, 8, 9) ~ "Other"))
# coverage status
mam_dat = mam_dat %>% 
  mutate(cover_cat  = case_when(NOTCOV == 1 | COVER == 4 | COVER65 == 6 ~ "None",
                                COVER == 2 | COVER65 %in% 2:4 ~ "Public",
                                COVER %in% c(1, 3) | COVER65 %in% c(1, 5) ~ "Private/Military"))

# disability
mam_dat = mam_dat %>% 
  mutate(lcond_chronic_cat = if_else(LCONDRT == 1, "Yes", "No"))

# race
mam_dat = mam_dat %>% 
  mutate(race_cat = case_when(RACRECI3 == 1 ~ "White",
                              RACRECI3 == 2 ~ "Black",
                              RACRECI3 == 3 ~ "Asian",
                              RACRECI3 == 4 ~ "AN/AI"),
         eth_cat = case_when(HISCODI3 == 1 ~ "Hispanic",
                             HISCODI3 == 2 ~ "Non-Hispanic White",
                             HISCODI3 == 3 ~ "Non-Hispanic Black",
                             HISCODI3 == 4 ~ "Non-Hispanic Asian",
                             HISCODI3 == 5 ~ "Non-Hispanic AN/AI"))


```


Survey Design
```{r}
mam_dat = mam_dat %>%
  mutate(domain = if_else(SEX == 2 & AGE_P >= 40, 1, 0))


des = svydesign(ids = ~PSU_P, strata = ~STRAT_P, weights = ~WTFA_SA, nest = TRUE, data = mam_dat)
```

Tables
```{r, tables}
# percent of women who have had mammogram in the last two years
# unsure-can't filter before the design because then it gives me an error about only one PSU in strata 73
# options(survey.lonely.psu = "certainty")
age_pct = svyby(~mam_2, by = ~domain+age_cat, svymean, na.rm = TRUE, design = des)
age_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
age_tot = svyby(~mam_2, by = ~domain+age_cat, svytotal, na.rm = TRUE, design = des)
age_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

# mammogram by education
edu_pct = svyby(~mam_2, by = ~domain+educ_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
edu_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
edu_tot = svyby(~mam_2, by = ~domain+educ_cat, svytotal, na.rm = TRUE, design = des)
edu_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

# mammogram by finc
finc_pct = svyby(~mam_2, by = ~domain+finc_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
finc_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
finc_tot = svyby(~mam_2, by = ~domain+finc_cat, svytotal, na.rm = TRUE, design = des)
finc_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

# mammogram by usual care
ausualp_pct = svyby(~mam_2, by = ~domain+ausualpl_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
ausualp_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
ausualp_tot = svyby(~mam_2, by = ~domain+ausualpl_cat, svytotal, na.rm = TRUE, design = des)
ausualp_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

# mammogram by health coverage
cover_pct = svyby(~mam_2, by = ~domain+cover_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
cover_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
cover_tot = svyby(~mam_2, by = ~domain+cover_cat, svytotal, na.rm = TRUE, design = des)
cover_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

# mammogram by chronic conditions
lcond_chronic_pct = svyby(~mam_2, by = ~domain+lcond_chronic_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
lcond_chronic_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
lcond_chronic_tot = svyby(~mam_2, by = ~domain+lcond_chronic_cat, svytotal, na.rm = TRUE, design = des)
lcond_chronic_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

#mammogram by race
race_pct = svyby(~mam_2, by = ~domain+race_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
race_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
race_tot = svyby(~mam_2, by = ~domain+race_cat, svytotal, na.rm = TRUE, design = des)
race_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()

# mammogram by ethnicity
eth_pct = svyby(~mam_2, by = ~domain+eth_cat, svymean, na.rm = TRUE, design = des, vartype = c("se", "ci"))
eth_pct %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
eth_tot = svyby(~mam_2, by = ~domain+eth_cat, svytotal, na.rm = TRUE, design = des)
eth_tot %>% filter(domain == 1) %>% select(-domain) %>% knitr::kable()
```


Tables By Age Group 

```{r, age_table}
# mammogram by education
edu_pct_strat = svyby(~mam_2, by = ~domain + age_cat + educ_cat, svymean, 
                      na.rm = TRUE, design = des, vartype = c("se", "ci"))
edu_tab = edu_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

edu_tab %>% knitr::kable()

edu_tot_strat = svyby(~mam_2, by = ~domain + age_cat + educ_cat, svytotal, 
                      na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)
edu_tot_strat %>% 
  knitr::kable()

num = edu_tot_strat$num

edu_tab = cbind(edu_tab, num) 

# mammogram by finc
finc_pct_strat = svyby(~mam_2, by = ~domain + age_cat + finc_cat, svymean, 
                       na.rm = TRUE, design = des, vartype = c("se", "ci"))

finc_tab = finc_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

finc_tab %>% knitr::kable()

finc_tot_strat = svyby(~mam_2, by = ~ domain + age_cat + finc_cat, svytotal,
                       na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = finc_tot_strat$num

finc_tab = cbind(finc_tab, num)


# mammogram by usual care
ausualp_pct_strat = svyby(~mam_2, by = ~domain + age_cat + ausualpl_cat,
                          svymean, na.rm = TRUE, design = des, 
                          vartype = c("se", "ci"))

usual_tab = ausualp_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

usual_tab %>% knitr::kable()

ausualp_tot_strat = svyby(~mam_2, by = ~domain + age_cat +ausualpl_cat,
                          svytotal, na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = ausualp_tot_strat$num

usual_tab = cbind(usual_tab, num)

# mammogram by health coverage
cover_pct_strat = svyby(~mam_2, by = ~domain + age_cat + cover_cat, svymean,
                        na.rm = TRUE, design = des, vartype = c("se", "ci"))

ins_tab = cover_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se)

ins_tab %>% knitr::kable()

cover_tot_strat = svyby(~mam_2, by = ~domain + age_cat + cover_cat, svytotal,
                        na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = cover_tot_strat$num

ins_tab = cbind(ins_tab, num)

# mammogram by chronic conditions
lcond_chronic_pct_strat = svyby(~mam_2, 
                                by = ~domain + age_cat + lcond_chronic_cat,
                                svymean, na.rm = TRUE, design = des,
                                vartype = c("se", "ci"))

dis_tab = lcond_chronic_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

dis_tab %>% knitr::kable()

lcond_chronic_tot_strat = svyby(~mam_2, 
                                by = ~domain + age_cat + lcond_chronic_cat,
                                svytotal, na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = lcond_chronic_tot_strat$num

dis_tab = cbind(dis_tab, num)

# mammogram by race
race_pct_strat = svyby(~mam_2, by = ~domain + age_cat + race_cat, svymean, 
                       na.rm = TRUE, design = des, vartype = c("se", "ci"))

race_tab = race_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

race_tab %>% knitr::kable()

race_tot_strat = svyby(~mam_2, by = ~domain + age_cat + race_cat, svytotal,
                       na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = race_tot_strat$num

race_tab = cbind(race_tab, num)

# mammogram by ethnicity
eth_pct_strat = svyby(~mam_2, by = ~domain + age_cat + eth_cat, svymean, 
                      na.rm = TRUE, design = des, vartype = c("se", "ci"))

eth_tab = eth_pct_strat %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

eth_tab %>% knitr::kable()

eth_tot_strat = svyby(~mam_2, by = ~domain + age_cat + eth_cat, svytotal, 
                      na.rm = TRUE, design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = eth_tot_strat$num

eth_tab = cbind(eth_tab, num)

# total gotten mammogram
total = svyby(~mam_2, by = ~domain + age_cat, svymean, na.rm = TRUE, 
              design = des, vartype = c("se", "ci"))

tot_tab = total %>% 
  filter(domain == 1) %>% 
  select(-domain, -se) 

tot_tab %>% knitr::kable()

tot_num = svyby(~mam_2, by = ~domain + age_cat, svytotal, na.rm = TRUE, 
                design = des) %>%
  filter(domain == 1) %>%
  select(-domain, -se) %>%
  rename(num = mam_2)

num = tot_num$num

tot_tab = cbind(tot_tab, num)
```

Combine into paper table

```{r}
tot_tab2 = tot_tab %>%
  mutate(type = "Total",
         level = "-")

edu_tab2 = edu_tab %>%
  mutate(type = "Education") %>%
  rename(level = educ_cat)

finc_tab2 = finc_tab %>%
  mutate(type = "Family Income Poverty Ratio") %>%
  rename(level = finc_cat)

usual_tab2 = usual_tab %>%
  mutate(type = "Usual Source of Care") %>%
  rename(level = ausualpl_cat)

ins_tab2 = ins_tab %>%
  mutate(type = "Insurance Type") %>%
  rename(level = cover_cat)

dis_tab2 = dis_tab %>%
  mutate(type = "Chronic Disability") %>%
  rename(level = lcond_chronic_cat)

eth_tab2 = eth_tab %>%
  mutate(type = "Ethnicity") %>%
  rename(level = eth_cat)

race_tab2 = race_tab %>%
  mutate(type = "Race") %>%
  rename(level = race_cat)

# create table of percentages of women who have gotten mammograms within the last two years (still need to add CIs)
tab_one = rbind(tot_tab2, edu_tab2, finc_tab2, usual_tab2, ins_tab2, dis_tab2, eth_tab2, race_tab2) %>%
  mutate(mam_2 = round(mam_2*100, 2),
         ci_l = round(ci_l*100, 2),
         ci_u = round(ci_u*100, 2),
         CI = str_c(ci_l, ", ", ci_u)) %>%
  rename(Percent = mam_2,
         No = num) %>%
  select(-ci_l, -ci_u) %>%
  pivot_wider(names_from = age_cat, values_from = c(No, Percent, CI)) %>%
  janitor::clean_names() %>%
  select(type, level, no_40_49, percent_40_49, ci_40_49, no_50_64, percent_50_64, ci_50_64, everything())

# print percentages
tab_one %>% knitr::kable()
```

Barplots of Usual Care

```{r}
b = svyby(~mam_2, by = ~domain + age_cat + ausualpl_cat,
                          svymean, na.rm = TRUE, design = des) %>%
  filter(domain == 1,
         ausualpl_cat != "Other") %>%
  select(-domain)

b1 = b %>%
  filter(trimws(age_cat) == "40–49")

b2 = b %>%
  filter(trimws(age_cat) == "50–64")

b3 = b %>%
  filter(age_cat == "65+")



barplot(b1$mam_2, beside = FALSE,
        names.arg = c("No", "Yes"),
        main = "Recent Mammogram for Women 40-49 by Usual Care",
        ylim = c(0, 1))

barplot(b2$mam_2, beside = FALSE,
        names.arg = c("No", "Yes"),
        main = "Recent Mammogram for Women 50-64 by Usual Care",
        ylim = c(0, 1))

barplot(b3$mam_2, beside = FALSE,
        names.arg = c("No", "Yes"),
        main = "Recent Mammogran for Women 65+ by Usual Care",
        ylim = c(0, 1))
```


Models

```{r}
mam_fit = svyglm(mam_2 ~ as.factor(age_cat) + as.factor(educ_cat) + as.factor(finc_cat) + as.factor(ausualpl_cat) + as.factor(cover_cat) + as.factor(lcond_chronic_cat) + as.factor(race_cat) + as.factor(eth_cat), 
       design = des, subset = domain == 1, family = binomial(link = "logit"))

summary(mam_fit)

Anova(mam_fit, type = 3)

mam_fit2 = svyglm(mam_2 ~ as.factor(age_cat) + as.factor(educ_cat) + as.factor(finc_cat) + as.factor(ausualpl_cat) + as.factor(cover_cat) + as.factor(eth_cat), design = des, subset = domain == 1,
                  family = binomial(link = "logit"))

summary(mam_fit2)

Anova(mam_fit2, type = 3)
```

Age cateogry, education level, financial status, usual care, insurance type, and ethnicity are significant predictors of having a recent mammogram.   
The OR for having a recent mammogram when comparing 50-64 year old women to 40-49 year old women is 0.9448.   
The OR for having a recent mammogram when comparing 65+ women to 40-49 year old women is 0.6784.   
The OR for having a recent mammogram when comparing women who only have a high school degree to women who have a college degree is 0.6985.   
The OR for having a recent mammogram when comparing women who have less than a high school degree to women who have a college degree is 0.5821.   
The OR for having a recent mammogram when comparing women who have some college education to women who have a college degree is 0.7943.   
The OR for having a recent mammogram when comparing women who are only known as having a poverty ratio of >200% to women who have a poverty ratio <200% is 1.1436.   
The OR for having a recent mammogram when comparing women who have a poverty ratio between 200-299% to women who have a poverty ratio <200% is 1.1586.   
The OR for having a recent mammogram when comparing women who have a poverty ratio between 300-399% to women who have a poverty ratio <200% is 1.4398.   
The OR for having a recent mammogram when comparing women who have a poverty ratio between 400-499% to women who have a poverty ratio <200% is 1.5428.   
The OR for having a recent mammogram when comparing women who have a poverty ratio >=500% to women who have a poverty ratio <200% is 1.8111.   
The OR for having a recent mammogram when comparing having a source of usual care to not having a source of usual care is 2.8878.   
The OR for having a recent mammogram when comparing having private/military insurance to none is 2.5213.   
The OR for having a recent mammogram when comparing having public insurance to none is 2.3978.   
The OR for having a recent mammogram when comparing Non-Hispanic AN/AI women to Hispanic women is 0.623.   
The OR for having a recent mammogram when comparing Non-Hispanic Asian women to Hispanic women is 0.4583.   
The OR for having a recent mammogram when comparing Non-Hispanic Black women to Hispanic women is 0.8474.   
The OR for having a recent mammogram when comparing Non-Hispanic White women to Hispanic women is 0.5621.   

The groups least likely to have had a recent mammogram are women aged 65+, women who have less than a high school degree, women who have a poverty ratio <200%, women who do not have a usual source of care, women who have no insurance, and Non-Hispanic women.